Models of gravitational lens candidates from Space Warps CFHTLS

### 2017-03-30 parts/figures.tex
AV: 2nd round of changes, updated table

\begin{figure*}
\includegraphics[width=\linewidth]{img/mlens_vs_mstel_one/mstel_vs_mtot_one}
\caption{Total mass in the model against the estimated stellar mass,
  alongside the values for the whole sample.  (The labelled orange
  bars are the systems shown in detail in
  Figures~\ref{fig:splinput}--\ref{fig:encl}.)  The lower-right shaded
  region is unphysical according to the stellar-population models,
  because it gives $M<\Mstel$. The upper-left shaded region is
  unphysical according to abundance matching (see
  Section~\ref{sec:stellar-mass}) because it gives $M>\Mhalo$.  That
  is to say, the unshaded region is
  $0<\haloindex<1$. \label{fig:stelmass}}
\end{figure*}

\begin{table*}
  \caption{Categorisation of SW models}
  \label{tab:models}
  \include{parts/table}
\end{table*}

>>>>

\begin{figure*}
\includegraphics[width=\linewidth]{img/mlens_vs_mstel_one/mstel_vs_mtot_one}
\caption{Total mass in the model against the estimated stellar mass,
  alongside the values for the whole sample.  (The labelled orange
  bars are the systems shown in detail in
  Figures~\ref{fig:splinput}--\ref{fig:encl}.)  The lower-right shaded
  region is unphysical according to the stellar-population models,
  because it gives $M<\Mstel$. The upper-left shaded region is
  unphysical according to abundance matching (see
  Section~\ref{sec:stellar-mass}) because it gives $M>\Mhalo$.  That
  is to say, the unshaded region is
  $0<\haloindex<1$.
  The width of the bars represent the error in $\Mstel$, where the
  lower and upper bounds are given assuming a ``young'' / ``old'' age trend
  for the population synthesis models (see Section~\ref{sec:stellar-mass}).
  The height of the bar is given by the minima and maxima values in $\Mhalo$
  obtained from the statistical ensemble of mass maps that make up
  one {\SpL} model (see Section~\ref{sec:massmodels}).
  \label{fig:stelmass}}
\end{figure*}

\begin{table*}
  \caption{Diagnostics of a selected model for each Space Warps candidate
    (see Section~\ref{sec:summary}).}
  \label{tab:models}
  \include{parts/table}
\end{table*}

### 2017-03-30 parts/intro.tex
AV: 2nd round of changes, updated table

The expected flood of new lens discoveries will need a similarly huge
modelling effort to reconstruct their mass distributions.  To prepare
for the challenge of large-sample lens modelling,
\cite{2015MNRAS.447.2170K} developed a new modelling strategy,
implemented as the SpaghettiLens system.  The idea is to collaborate
with experienced members of the citizen-science community, who have
already participated in lens discovery through {\SW}, as well as
several other projects involving astronomical data.  The system was
tested on a sample of simulated lenses generated as part of the
training set used in {\SW} (see
\cite{2016MNRAS.455.1191M,2016MNRAS.455.1171M} for further information).

@@

\begin{itemize}
\item First we have diagnostics based on morphology of the
  system.
  Section~\ref{sec:morph} and Figure~\ref{fig:splinput} explain.
\begin{itemize}
\item Whether the images are unblended.  Distinct unblended images are
  an advantage in modelling, but not essential.
\item Whether all images are discernible.  The topography of an
  arrival-time surface, as encoded by a spaghetti diagram, may require
  more images than are visible, in which case the modeller has to
  insert conjectural image positions.
\item Whether the lens is fairly isolated.
\item The image morphology concisely described: double or quads,
  further sub-categorised to indicate the elongation direction of the
  lensing mass.
\end{itemize}
\item Second we have mass models, covered in Section~\ref{sec:massmodels}.
\begin{itemize}
\item Whether the mass map is reasonable. Figure~\ref{fig:kappa}.
\item Whether the arrival-time surface and synthetic image are
  plausible.  In particular, additional images are implied in regions
  where they are not observed signal an unsatisfactory model.
  Figures~\ref{fig:arriv} and \ref{fig:synth} and \ref{fig:encl}.
\end{itemize}
\item Third, whether the implied lensing mass is plausible, given the
  photometry data of the lensing galaxy.  Section~\ref{sec:stellar-mass}
  explains how we compare the lensing mass with the mass in stars in
  the lensing galaxy.  We estimate the stellar mass by comparing
  galaxy magnitudes from the CFHTLS pipeline with the well-known
  stellar-population models of \cite{2003MNRAS.344.1000B}.  We then
  extrapolate the stellar mass to a halo mass using the
  abundance-matching prescription of \cite{2010ApJ...710..903M}.
  Naturally, the lensing mass must be more than the stellar mass but
  no more than the total halo mass.  We then introduce what we call a
  halo index ($\haloindex$) which gives an idea of how the lensing mass
  compares with these two constraints.  Figure~\ref{fig:stelmass}.
\end{itemize}

Section~\ref{sec:summary} summarises and tabulates the diagnostics in
Table~\ref{tab:models}.  Interpretation of the results is tentative,
because the systems are lens candidates at this stage, not secure lenses.
Moreover the candidate-lens redshifts have large uncertainties, while
the candidate-source redshifts can only be guessed at present.
Nevertheless it is interesting to see what trends we can observe with
the already-available data.

There are three appendices devoted to various technical issues
relating to modelling.  \cite{2015MNRAS.447.2170K} tested the system
on simulated lenses and identified some areas for improvement.  In
\S~\ref{subsec:sourcefit} we introduce fitting of the brightness
profiles of the source.  This feature has not yet been included in
SpaghettiLens, but has been carried out in post-processing for a few
especially interesting candidates.  In \S~\ref{subsec:hires} we show
that making high resolution mass maps in the central region relieves a
tendency in the earlier work for mass to be too shallow. Then in
\S~\ref{subsec:parameter} we consider the possibility of fitting a
parametric lens model to the model ensemble; so far we have only been
successful at extracting an Einstein radius.

>>>>

The expected flood of new lens discoveries will need a similarly huge
modelling effort to reconstruct their mass distributions.  To prepare
for the challenge of massive-sample lens modelling,
\cite{2015MNRAS.447.2170K} developed a new modelling strategy,
implemented as the SpaghettiLens system.  The idea is to collaborate
with experienced members of the citizen-science community, who have
already participated in lens discovery through {\SW}, as well as
several other projects involving astronomical data.  The system was
tested on a sample of simulated lenses\footnote{
  Those simulated lenses were generated as part of the training set used
  in {\SW} (see \cite{2016MNRAS.455.1191M,2016MNRAS.455.1171M} for further
  information)}
from \SW.

@@

\begin{itemize}
\item First we have diagnostics based on morphology of the
  system.
  Section~\ref{sec:morph} and Figure~\ref{fig:splinput} explain.
\begin{itemize}
\item Whether the images are unblended.  Distinct unblended images are
  an advantage in modelling, but not essential.
\item Whether all images are discernible.  The topography of an
  arrival-time surface, as encoded by a spaghetti diagram, may require
  more images than are visible, in which case the modeller has to
  insert conjectural image positions.
\item Whether the lens is fairly isolated.
\item The image morphology concisely described: double or quads,
  further sub-categorised to indicate the elongation direction of the
  lensing mass.
\end{itemize}
\item Second we have mass models, covered in Section~\ref{sec:massmodels}.
\begin{itemize}
\item Whether the mass map is reasonable. Figure~\ref{fig:kappa}.
\item Whether the arrival-time surface and synthetic image are
  plausible.  In particular, additional images are implied in regions
  where they are not observed signal an unsatisfactory model.
  Figures~\ref{fig:arriv} and \ref{fig:synth} and \ref{fig:encl}.
\end{itemize}
\item Third, whether the implied lensing mass is plausible, given the
  photometry of the lensing galaxy.  Section~\ref{sec:stellar-mass}
  explains how we compare the lensing mass with the mass in stars in
  the lensing galaxy.  We estimate the stellar mass by comparing
  galaxy magnitudes from the CFHTLS pipeline with the well-known
  stellar-population models of \cite{2003MNRAS.344.1000B}.  We then
  extrapolate the stellar mass to a halo mass using the
  abundance-matching prescription of \cite{2010ApJ...710..903M}.
  Naturally, the lensing mass must be more than the stellar mass but
  no more than the total halo mass.  We then introduce what we call a
  halo index ($\haloindex$) which gives an idea of how the lensing mass
  compares with these two constraints.  Figure~\ref{fig:stelmass}.
\end{itemize}

Section~\ref{sec:summary} summarises and tabulates the diagnostics in
Table~\ref{tab:models}.  Interpretation of the results is tentative,
because the systems are lens candidates at this stage, not secure lenses.
Moreover the candidate-lens redshifts have large uncertainties, while
the candidate-source redshifts can only be guessed at present.
Nevertheless it is interesting to see what trends we can observe with
the already-available data.
%TODO there is more info about uncertainies here:
% https://arxiv.org/pdf/0811.3326.pdf

There are three appendices devoted to various technical issues
relating to modelling.  \cite{2015MNRAS.447.2170K} tested the system
on simulated lenses and identified some areas for improvement.  In
\S~\ref{subsec:sourcefit} we introduce fitting of the brightness
profiles of the source.  This feature has not yet been included in
SpaghettiLens, but has been carried out in post-processing for a few
especially interesting candidates.  In \S~\ref{subsec:hires} we show
that making mass maps fine-grained in the central region relieves a
tendency in the earlier work for mass profile to be too shallow. Then in
\S~\ref{subsec:parameter} we consider the possibility of fitting a
parametric lens model to the model ensemble; so far we have only been
successful at extracting an Einstein radius.

### 2017-03-30 parts/massmodel.tex
AV: 2nd round of changes, updated table

The mass will naturally depend on the lens and source redshifts, which
are unknown when a lens candidate is first identified.  But this is
not a problem, because a model can be trivially rescaled to use better
redshift values, as and when they become available.  This work applies
photometric redshifts from the CFHTLS pipeline to the candidate
lensing galaxies; the values range from $z=0.2$ to $z=1$ (see
Table~\ref{tab:models}).  Source redshifts are set to $z=2$ unless an
unambiguous photometric redshift is available.  The mass normalisation
depends only weakly on the source redshift, provided it is not close
to the lens redshift.

>>>>

The mass will naturally depend on the lens and source redshifts, which
are unknown when a lens candidate is first identified.  But this is
not a problem, because a model can be trivially rescaled to use better
redshift values, as and when they become available.  This work applies
photometric redshifts from the CFHTLS pipeline to the candidate
lensing galaxies; the values range from $z=0.2$ to $z=1$ (see
Table~\ref{tab:models}).  As the mass normalisation depends only weakly
on source redshift, provided it is larger than to the lens redshift,
we assumed all sources to lie at $z=2$ unless an unambiguous photometric
redshift is available.

### 2017-03-30 parts/morph.tex
AV: 2nd round of changes, updated table

A citizen modeller's main input to the lens-modelling process is a markup of
the candidate lens system, which we call a spaghetti diagram.
This is a sketch of the arrival-time surface from a point-like source,
with user-proposed locations of maxima, minima and saddle points, and an
implied time-ordering of the images.  Such information encodes a
starting proposal for the mass distribution.  A spaghetti diagram is
thus a completely abstract construction, and moreover it refers to a
simplified system with a point source.  However, spaghetti diagrams
are intuitive because they tend to resemble the form of lensed arcs,
and of course they are simple to draw, and easy to vary and refine in
an open collaborative environment.  This makes them very practical for
non-professional lens enthusiasts in the citizen-science community.
Details and tests are given in \cite{2015MNRAS.447.2170K}.

>>>>

A modeller's main input to the lens-modelling process is a markup of
the candidate lens system, which we call a spaghetti diagram.
This is a sketch of the arrival-time surface from a point-like source,
with proposed locations of maxima, minima and saddle points, and an
implied time-ordering of the images.  Such information encodes a
starting proposal for the mass distribution.  A spaghetti diagram is
thus a completely abstract construction, and moreover it refers to a
simplified system with a point source.  However, spaghetti diagrams
are intuitive because they tend to resemble the form of lensed arcs,
and of course they are simple to draw, and easy to vary and refine in
an open collaborative environment.  This makes them very practical for
non-professional lens enthusiasts in the citizen-science community.
Details and tests are given in \cite{2015MNRAS.447.2170K}.
